---
output:
  html_document: default
  pdf_document: default
---
```{r knitr_setup, echo=FALSE}
knitr::opts_chunk$set(cache=TRUE)
```

## Data types

_1) Provide a URL to the dataset._

I downloaded my dataset from https://www.kaggle.com/tmdb/tmdb-movie-metadata#tmdb_5000_movies.csv


_2) Explain why you chose this dataset._

I enjoy watching movies and this was one of the first data sets that caught my eye


_3) What are the entities in this dataset? How many are there?_

The entities in this dataset are top movies from TMDb. There are 4803 entities

_4) How many attributes are there in this dataset?_

There are 20 attributes

_5) What is the datatype of each attribute (categorical -ordered or unordered-, numeric -discrete or continuous-, datetime, geolocation, other)? Write a short sentence stating how you determined the type of each attribute. Do this for at least 5 attributes, if your dataset contains more than 10 attributes, choose 10 of them to describe._

| Num | Name | Type | Description |
|-----|------|------|-------------|
| 1 | `original_title` | other - title | Title of the movie |
| 2 | `tagline` | other - tagline | Tagline of the movie |
| 3 | `budget` | numeric continuous | total budget of the movie |
| 4 | `status` | categorical unordered | Can take value from finite set of possible statuses |
| 5 | `vote_count` | numeric continuous | Total count of votes |
| 6 | `vote_average` | numeric discrete | Average vote on a scale of 1-10 |
| 7 | `release_date` | datetime | Specifies date of release |
| 8 | `revenue` | numeric continous | Total revenue earned |
| 9 | `runtime` | other - address | Stree address if incident |
| 10 | `original_language` | categorical unordered | Can take value from finite set of possible languages |

_6) Write R code that loads the dataset using function `read_csv`. Were you able to load the data successfully? If no, why not?_

```{r load_data, message=FALSE}
library(tidyverse)
#When i first used the url provided above, I was receiving a parsing error due to some sort of format error. However, once I uploaded it from my local machine it worked perfectly
movies <- read_csv('tmdb_5000_movies.csv')

```

## Wrangling

1) My pipeline computes the average budget by original language (ignores budgets <=0)

```{r pipeline}
mean_budgets <- movies %>%
  filter(budget > 0) %>%
  select(original_language, budget) %>%
  group_by(original_language) %>%
  summarize(mean_budget=mean(budget)) %>%
  arrange(mean_budget)
mean_budgets
```

## Plotting

1) This barplot shows the average budget per original_language (ignoring budgets <= 0)

```{r plot}
mean_budgets %>%
  ggplot(aes(x=original_language, y=mean_budget)) +
    geom_bar(stat="identity") +
    coord_flip()
```