---
title: "Interactive Map of Baltimore"
author: "Daniel Brewer"
date: "5/7/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

A dataset containing Baltimore Crime statistics was used to create an interactive map of Baltimore Crime.
```{r}

library(tidyverse)
library(stringr)

arrest_tab <- read_csv("http://www.hcbravo.org/IntroDataSci/misc/BPD_Arrests.csv")
dat <- arrest_tab %>% 
  filter(!is.na(`Location 1`)) %>%
  separate(`Location 1`, c("lat","lng"), sep=",") %>%
  mutate(lat=as.numeric(str_replace(lat, "\\(", ""))) %>%
  mutate(lng=as.numeric(str_replace(lng, "\\)", ""))) %>%
  sample_n(2000)
dat

```

In this interactive map, arrest incidents are shown using geographic data obtained from the data set. Each offense location is shown on the map using a violet or blue colored circle. A violet colored circle indicates that the arrest was made on a Female, while a light blue circle indicates the arrest was made on a Male. When a circle is clicked on, a pop-up is shown detailing the type of offense. Overall, the map is showing an interactive distribution of arrests categorized by gender.
```{r}
library(leaflet)
library(htmltools)

test <- colorFactor(palette = c("violet", "blue"), dat$sex)

balto_map <- leaflet(dat) %>% addTiles() %>% addCircleMarkers(
  radius = ~ifelse(sex == "M", 6, 8),
  color = ~test(sex), 
  stroke = FALSE, 
  fillOpacity = ~ifelse(sex == "M", 0.6, 1),
  popup = ~htmlEscape(incidentOffense)) %>% 
  
  addLegend(position = 'topright', pal = test, values = ~sex, labels = "Sex") %>%
  setView(lat=39.29, lng=-76.61, zoom=11)
balto_map

```
